---
title: "HW 18"
author: "Kevin H. Bhimani"
date: "May 27, 2018"
output:
  html_document: default
  pdf_document: default
---

Chapter 7: 26(4 points), 36 (4 points), 38

26. The data set MnGroundwater has measurements on water quality of 895 randomly selected wells in Minnesota.
(a) Create a histogram or normal quantile plot of the alkalinity and comment on the distribution.
```{r}
Minnesota <- read.csv("http://people.kzoo.edu/enordmoe/math365/data/MnGroundwater.csv")
alkalinity <- Minnesota$Alkalinity
hist(alkalinity)
qqnorm(alkalinity)
qqline(alkalinity)
```
The distriubiton looks even distributed. The normal qq plot confirms that since we have almost a straight line.

(b) Find a 95% t confidence interval for the mean $\mu$ of alkalinity levels in Minnesota wells.
we set $\alpha = 0.05$. Then confidence interval is = $(\bar{X}-q  \frac{S}{\sqrt{n}},\bar{X}+q \frac{\sigma}{\sqrt{n}})$

$q=qnorm(1-0.05/2) = 1.959964$
```{r}
t.test(alkalinity,conf.level=0.95)$conf
```
So the confidence interval is (283575.6,297789.8)

(c) Find the 95% bootstrap percentile and bootstrap t confidence intervals for the mean alkalinity level and compare results. Which confidence interval will you report?
```{r}
Minnesota <- read.csv("http://people.kzoo.edu/enordmoe/math365/data/MnGroundwater.csv")
alkalinity <- Minnesota$Alkalinity
xbar <- mean(alkalinity)
N <- 10^4
n <- length(alkalinity)
mean.bootstrap <- numeric(N)
Tstar <- numeric(N)
set.seed(365)
for (i in 1:N)
{
x <- sample(alkalinity, size = n, replace = T)
mean.bootstrap[i] <- mean(x)
Tstar[i] <- (mean(x)-xbar) / (sd(x)/sqrt(n))
}
quantile(mean.bootstrap,c(.025, .975))
quantile(Tstar, c(.975,.025))
 q1 <- -1.945211
 q2 <- 1.928312 
 xbar - q2*(sd(alkalinity)/sqrt(length(alkalinity)))
 xbar - q1*(sd(alkalinity)/sqrt(length(alkalinity)))

```
95% bootstrap percentile: (283648.0,297592.3 )
95% bootstrap t confidence interval:(283513.8,297891.4)

Because of large sample size, I would report 95% bootstrap t confidence interval
```{r}
xbar <- mean(alkalinity)
xbar
xbar - 1.96*(sd(alkalinity)/sqrt(895))
xbar + 1.96*(sd(alkalinity)/sqrt(895))
```

95% bootstrap t confidence interval: (283585.1,297780.3)

Because of the large sample size, I would report the

36. In this exercise, you will derive and use a "bootstrap Z" interval.
(a) Following the steps in the derivation of the bootstrap t interval in Section 7.5,
derive a bootstrap Z interval for $\mu$, for cases when $\sigma$ is known.

We can replace S with $\sigma$ for formula in Pg. 197
$(\bar{x}-Q_2^*\frac{\sigma}{\sqrt{n}},\bar{x}-Q_1^*\frac{\sigma}{\sqrt{n})})$

(b) Calculate this interval for the Verizon CLEC data; for $\sigma$, use the sample
variance of the Verizon ILEC data. (In practice, we may use methods for known $\sigma$ when we can estimate ?? from a large related data set.)

```{r}
Verizon <- read.csv("http://people.kzoo.edu/enordmoe/math365/data/Verizon.csv")
ILEC <- subset(Verizon, select="Time", Group =="ILEC", drop=T)
CLEC <- subset(Verizon, select="Time", Group =="CLEC", drop=T)

xbar <- mean(CLEC)
sd <- sd(ILEC)
t.test(ILEC, conf.level = .95)$conf
N <- 10^4
n <- length(CLEC)
Zstar <- numeric(N)
set.seed(365)
for (i in 1:N)
{
x <- sample(CLEC, size = n, replace = T)
Zstar[i] <- (mean(x)-xbar)/(sd/sqrt(n))
}
 quantile(Zstar, c(0.025, 0.975))
```

```{r}
 q1 <- -2.064709
 q2 <- 2.875568 
 xbar - q2*(sd/sqrt(length(CLEC)))
 xbar - q1*(sd/sqrt(length(CLEC)))
```
So the confidence interval is (7.701021,22.83351)

(c) Compare that interval with a formula z interval. How does the bootstrap Z interval adjust for skewness?
```{r}
 xbar - 1.96*(sd/sqrt(length(CLEC)))
 xbar + 1.96*(sd/sqrt(length(CLEC)))
```

The Z confidence interval is (10.50548,22.51278)

The bootstrap Z interval does not assume normal distribution and thus adjust for skewness.

38. Robin wonders how much variation there is in a box of his favorite cereal. He buys eight boxes from eight different stores and finds the weights of the contents
to be (in grams)
560 568 580 550 581 581 562 550
Assuming the data are from a normal distribution, find a 90%confidence interval for the variance $\sigma^2$. (In R, the command qchisq computes quantiles for the
chi-square distribution.)
theorem b16

```{r}
v <- c(560, 568, 580, 550, 581, 581, 562, 550)
xbar <- mean(v)
xvar <- var(v)
n <-8
a <- qchisq(0.05,7)
b <- qchisq(0.95,7)
((n-1)*xvar)/b
((n-1)*xvar)/a


```
So the interval is (86.15824,559.2083)
