---
title: "Math Stats HW #3"
subtitle: "Handin Problems:  2.12, 2.14, 2.16"
author: "Kevin Bhimani"
date: "April 11, 2018"
output: html_document
---

```{r setup, include=FALSE}
# This chunk sets a couple options to keep messages and warnings
# from appearing in the final write-up.
# It also sets default rounding to 3 decimal places.
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning=FALSE)
options(digits = 4,width=100,fig.width=8) 
```



#### Problem 2.12

Let $X$ be a random variable with cdf $F(x) = 1-9/x^2$ for $x\ge 3$. Find an expression for the $p$th quantile of $X$.

**Answer:**
 We know $F(x)= 1-9/x^2$. Then for $p$th quantile,
 $1-9/x^2 = p$
 or $1+p = 9/x^2$
 or $x = \pm \frac{3}{\sqrt(1+p)}$
 


#### Problem 2.14

In this exercise, we investigate normal quantile plots using R.  

(a) Draw a random sample of size $n = 15$ from $N(0,1)$ and plot both the normal quantile plot and the histogram. Do the points on the quantile plot appear to fall on a straight line? Is the histogram symmetric, unimodal, and mound shaped? Do this several times.  

**Answer:**

```{r}
x <- rnorm(15) # draw random sample of size 15 from N(0,1)
#par(mfrow=c(2,1)) # set up plot area to place 2 graphs on one sheet
qqnorm(x)
qqline(x)
hist(x)

```

(b) Repeat part (a) for samples of size $n = 30$, $n = 60$, and $n = 100$.   

**Answer:**

```{r}
x <- rnorm(30) # draw random sample of size 15 from N(0,1)
#par(mfrow=c(2,1)) # set up plot area to place 2 graphs on one sheet
qqnorm(x)
qqline(x)
hist(x)

x <- rnorm(60) # draw random sample of size 15 from N(0,1)
#par(mfrow=c(2,1)) # set up plot area to place 2 graphs on one sheet
qqnorm(x)
qqline(x)
hist(x)

x <- rnorm(100) # draw random sample of size 15 from N(0,1)
#par(mfrow=c(2,1)) # set up plot area to place 2 graphs on one sheet
qqnorm(x)
qqline(x)
hist(x)

```


(c) What lesson do you draw about using graphs to assess whether or not a data set follows lesson a normal distribution?  

We see that the data does follow a normal distribution since we are getting a straight line. It is evident that the normal behavior increases once the sample size is increased.

**Answer:**


#### Problem 2.16

The ecdf for a data set with $n=20$ values is given in Figure 2.16.  

(a) How many values are less than or equal to 7?   

**Answer:**  
3

(b) How many times does the value 8 occur?   

**Answer:**  
4

(c) In a histogram of these values, how many values fall in the bin (20,25]?  

**Answer:**  
6
