---
title: "Math Stats HW #4"
subtitle: "Handin Problems:  3.4 (4 points), 3.6 (4 points), 3.8 (2 points)"
author: "Kevin H. Bhimani"
date: "April 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
# This chunk sets a couple options to keep messages and warnings
# from appearing in the final write-up.
# It also sets default rounding to 3 decimal places.
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning=FALSE)
options(digits = 4,width=100,fig.width=8) 
```


```{r, include=FALSE}
# These commands silently read in the datsets without producing 
# output
FlightDelays<- read.csv("http://people.kzoo.edu/enordmoe/math365/data/FlightDelays.csv")
Spruce<-read.csv("http://people.kzoo.edu/enordmoe/math365/data/Spruce.csv")
```


#### Problem 3.4  

In the **FlightDelays** Case Study in Section 1.1, the data contain flight delays for two airlines, American Airlines and United Airlines.

(a) Compute the proportion of times that each carrier's flights was delayed more than 20 min. Conduct a two-sided test to see if the difference in these proportions is statistically significant.   

**Answer:**  

```{r}
Del<-FlightDelays$Delay
Del.UA<-subset(FlightDelays, select=Delay, subset=Carrier=="UA", drop=T)
Del.AA<-subset(FlightDelays, select=Delay, subset=Carrier=="AA", drop=T)
u <- mean(Del.UA>20)
v <- mean(Del.AA>20)
z <- u - v
z
```
```{r}
N<-10^4-1
reslt<-numeric(N)
for(i in 1:N)
{
index<-sample(4029, size=1123, replace=FALSE)
reslt[i]<-mean(Del[index]>20)-mean(Del[-index]>20)
}
pValue<-(sum(reslt>=z)+1)/(N+1)
pValue*2
```


(b) Compute the variance in the flight delay lengths for each carrier. Conduct a test to see if the variance for United Airlines is greater than that of American Airlines.     

**Answer:**  

```{r}
var(Del.UA)
var(Del.AA)
a <- var(Del.UA)/var(Del.AA)
a

N<-10^4-1
reslt<-numeric(N)
for(i in 1:N)
{
index<-sample(4029, size=1123, replace=FALSE)
reslt[i]<-mean(Del[index]>20)-mean(Del[-index]>20)
}
pValue<-(sum(reslt>=a)+1)/(N+1)
pValue*2
```

```




#### Problem 3.6  

In the Flight Delays Case Study in Section 1.1,   

(a) Find the trimmed mean of the delay times for United Airlines and American Airlines.  

**Answer:**  


```{r}
mean(Del.UA, trim = .25)
mean(Del.AA, trim = .25)

```


(b) Conduct a two-sided test to see if the difference in trimmed means is statistically significant.  

**Answer:**  


```{r}
t <- mean(Del.UA, trim = .25)-mean(Del.AA, trim = .25)
t

N<-10^4-1
reslt<-numeric(N)
for(i in 1:N)
{
index<-sample(4029, size=1123, replace=FALSE)
reslt[i]<-mean(Del[index], trim = .25)-mean(Del[-index], trim = .25)
}
pValue<-(sum(reslt>=t+1))/(N+1)
2*pValue
```


#### Problem 3.8  

In the Black Spruce Case Study in Section 1.9, seedlings were planted in plots that were either subject to competition (from other plants) or not. Use the data set **Spruce** to conduct a test to see if the mean difference in how much the seedlings grew over the course of the study under these two treatments is statistically significant. 


```{r}
Htchange<-subset(Spruce, select = Ht.change, drop = TRUE)
Htchange.C<-subset(Spruce, select = Ht.change, subset = (Competition == "C"), drop = TRUE)
Htchange.NC<-subset(Spruce, select = Ht.change, subset = (Competition == "NC"), drop = TRUE)
observed<-mean(Htchange.NC)-mean(Htchange.C)
observed

N<-10^4-1
result<-numeric(N)
for(i in 1:N)
{
index<-sample(72, size=36, replace=FALSE)
result[i]<-mean(Htchange[index])-mean(Htchange[-index])
}
pValue<-(sum(result>=observed+1))/(N+1)
2*pValue
```

