---
title: "HW 8"
author: "Kevin H. Bhimani"
date: "April 24, 2018"
output: html_document
---
Q. 2
a)
```{r}
sample21 <- c(1,3,4,6)
N <- 10^5
my.boot <- numeric(N)
for (i in 1:N)
{
x <- sample(sample21, 1, replace = TRUE) # draw resample
my.boot[i] <- mean(x) # compute mean, store in my.boot
}

simList <- numeric(N)
for(i in 1:N){
 x <- my.boot[i]
 simList[i] <- 0 
 if (x==1){
   simList[i] <- 1
 }
}
mean(simList)
```
b)
```{r}
mean(my.boot < 6)
```

c)
Total favorable combinations = 54
Thus P = 54/256


Q. 8 Consider a population that has a gamma distribution with parameters r = 5, ?? = 1/4.

(a) Use simulation to generate an approximate sampling distribution of the
mean; plot and describe the distribution.
```{r}
sample <- rgamma(10^4,5,1/4)
mean(sample)
var(sample)
hist(sample)
```
The distribution looks skewed towards left.

(b) Now, draw one random sample of size 200 from this population. Create a
histogram of your sample and find the mean and standard deviation.
```{r}
sample2 <- rgamma(200,5,1/4)
mean(sample2)
sd(sample2)
hist(sample2)
```
(c) Compute the bootstrap distribution of the mean for your sample, plot it, and note the bootstrap mean and standard error.

```{r}
N <- 10^5
my.boot <- numeric(N)
for (i in 1:N)
{
x <- sample(sample2, 200, replace = TRUE) # draw resample
my.boot[i] <- mean(x) # compute mean, store in my.boot
}
mean(my.boot)
sd(my.boot)
hist(my.boot)
```

(d) Compare the bootstrap distribution to the approximate theoretical sampling
distribution by creating a table like Table 5.2.




(e) Repeat for sample sizes of n = 50 and n = 10. Describe carefully your
observations about the effects of sample size on the bootstrap distribution.

```{r}
N <- 10^5
my.boot <- numeric(N)
for (i in 1:N)
{
x <- sample(sample2, 50, replace = TRUE) # draw resample
my.boot[i] <- mean(x) # compute mean, store in my.boot
}
mean(my.boot)
sd(my.boot)
hist(my.boot)

N <- 10^5
my.boot <- numeric(N)
for (i in 1:N)
{
x <- sample(sample2, 10, replace = TRUE) # draw resample
my.boot[i] <- mean(x) # compute mean, store in my.boot
}
mean(my.boot)
sd(my.boot)
hist(my.boot)
```
I observe that the mean and standard deviation increase with decrease in sample size. I also see that the distribution is getting more skewed towards right.
