---
title: "HW 9"
author: "Kevin H. Bhimani"
date: "April 25, 2018"
output: html_document
---

10. Import the data from data set Bangladesh. In addition to arsenic concentrations for 271 wells, the data set contains cobalt and chlorine concentrations.

(a) Conduct EDA on the chlorine concentrations and describe the salient
features.

```{r}

Bangladesh <- read.csv("http://people.kzoo.edu/enordmoe/math365/data/Bangladesh.csv",header = TRUE)
chlorine <- subset(Bangladesh, select = Chlorine, subset =!is.na(Chlorine), drop = T)
hist(chlorine)
mean(chlorine)
var(chlorine)
```
(b) Bootstrap the mean.

```{r}
n <- length(chlorine)
t <-10^4
boot.mean <-  numeric(t)
for(i in 1:10^4){
x <- sample(chlorine,n,replace=TRUE)
boot.mean[i] <-mean(x)
}
mean(boot.mean)
var(boot.mean)
hist(boot.mean)
```

(c) Find and interpret the 95% bootstrap percentile confidence interval.
```{r}
quantile(boot.mean, c(0.025,0.975))
```

(d) What is the bootstrap estimate of the bias? What fraction of the bootstrap standard error does it represent?
```{r}
mean(boot.mean) - mean(chlorine) #bias
sd(boot.mean) #standard error
```
12. The data set FishMercury contains mercury levels (parts per million) for 30 fish caught in lakes in Minnesota.
(a) Create a histogram or boxplot of the data. What do you observe?
```{r}
FishMercury <- read.csv("http://people.kzoo.edu/enordmoe/math365/data/FishMercury.csv",header = TRUE)
Fish <- subset(FishMercury, select = Mercury , drop=T)
ca <- seq(0,2,0.25)
hist(Fish, ca)
```
This shows most of the data lies between 0 and 0.25

(b) Bootstrap the mean and record the bootstrap standard error and the 95% bootstrap percentile interval.
```{r}
n <- length(Fish)
B <- 10^4
fish.boot <- numeric(B)
for( i in 1:B){
  x <- sample(Fish,n,replace=TRUE)
  fish.boot[i] <- mean(x)
}
hist(fish.boot)
mean(fish.boot)
sd(fish.boot)
quantile(fish.boot, c(0.025,0.975))

```

(c) Remove the outlier and bootstrap the mean of the remaining data. Record the bootstrap standard error and the 95% bootstrap percentile interval.
```{r}
mercury <- FishMercury < max(FishMercury)
n <- length(Fish)-1
B <- 10^4
fish.boot1 <- numeric(B)
for( i in 1:B){
  x <- sample(mercury,n,replace=TRUE)
  fish.boot1[i] <- mean(x)
}
hist(fish.boot1)
mean(fish.boot1)
sd(fish.boot1)
quantile(fish.boot1, c(0.025,0.975))
```

(d) What effect did removing the outlier have on the bootstrap distribution, in particular, the standard error?

There was a big change in distribution. It became skewed towards right instead of left. The mean increased too and the standard deviation decreased.

